Growing up with Nell: A Narrative Interface for Literacy

### 2012-03-11 paper.tex
Tighten a few Technology sentences.

\section{A Lesson from Nell}
\begin{figure}
\centering
\epsfig{file=xo3-3.eps, width=3.2in}
\caption{Nell's solar-powered hardware platform.  The tablet design
  allows for direct interaction.}\label{fig:xo3}
\end{figure}
% tell a nell story up front
Miles from the nearest school, a young Ethiopian girl named Rahel turns on her
new tablet computer. The solar-powered machine speaks to her:
\textit{``Hello!  Would you like to hear a story?''}

@@

\section{Technology}
We have begun implementation of the Nell system.  In this section we
will describe the implementation of three of the most novel pieces
from the software engineering side.  These are not necessarily the
most novel-appearing from the user's perspective: for example, the
handwriting UI is still fairly unusual in consumer devices, but the
implementation of a separated-character handwriting recognition engine
using Hidden Markov Models is well-understood~\cite{hanshu:handwriting}.
\textbf{SMOOTH THIS LANGUAGE}

\subsection{Rule-based story modules}
\begin{figure}\small
\begin{verbatim}
{ world: { // id gives canonical URL for this module
    id: "/nell.laptop.org/chapters/alphabet-book",
    scene: "opening", // initial scene
    inherit_from: "/nell.laptop.org/core/drawing_activity",
  }, scenes: [
    { id: "opening", subject: "actors/princess",
      verb: "speak", object: "actors/user",
      text: "Hello! Would (o/2) like to hear a story?",
      group: "/nell.laptop.org/core/story-intro-group",
    },
    { id: "story-1",
      cond: rule`At(opening) && Event(Chat, Yes)`,
      subject: "actors/princess",
      verb: "speak", object: "actors/user",
      text: "Once upon a time..." },
    ...
    { id: "r-drawn",
      cond: rule`Event(Activity, DrawR)`,
      subject: "actors/princess",
      verb: "speak", object: "actors/user",
      text: "Very good, (o/3)!",
      action: js`startAnimation("r-runs")` }, ...
  ], ... }
\end{verbatim}
\caption{Excerpt of story module.  JSON syntax has been extended with
quasiquote~\cite{quasiquote}.}\label{fig:rules}
\end{figure}
Narrative is the core of the Nell interface.  It is important that
Nell's story be modular and extensible.  We use a rule-based story
model with programmable conflict resolution to provide this ability.
Figure~\ref{fig:rules} shows an excerpt of the story module driving
the interaction described in the introduction to this paper.

Stories consist of a sequence of \textit{scenes}, in addition to
characters, locations, and activity information.  Scenes are ``goals''
in the STRIPS formalism~\cite{strips}, and consist of
action descriptions, preconditions, and effects.
In principle, all preconditions for all scenes of all story modules
are evaluated continuously to determine the actions of the narrative
agents.  This production rules system
is made computationally efficient using the Rete
algorithm.  Dialog is expressed using Curveship-style
templates~\cite{montfort:curveship}.

@@

New story modules can be incorporated by visiting an internet page in
connected deployments.  In disconnected deployments, new stories
might be distributed once a year on USB sticks, or shared from your
friends.  In a classroom environment, a teacher can share stories for
the day's lesson at the start of class.

>>>>

\section{A Lesson from Nell} \label{sec:lesson}
\begin{figure}
\centering
\epsfig{file=xo3-3.eps, width=3.2in}
\caption{Nell's solar-powered hardware platform.  The tablet design
  allows for direct interaction.}\label{fig:xo3}
\end{figure}
% tell a nell story up front
Miles from the nearest school, a young Ethiopian girl named Rahel turns on her
new tablet computer. The solar-powered machine speaks to her:
\textit{``Hello!  Would you like to hear a story?''}

@@

\section{Technology}
We describe the novel synthesis of technologies that makes up our current Nell
prototype. Several of these technologies are well-established; a few (like
``TurtleScript'') are novel.

\subsection{Handwriting recognition}
To support \textit{direct interaction}, we use a separated-character
handwriting recognition engine based on Hidden Markov Models~\cite{hanshu:handwriting}.

\subsection{Rule-based story modules}
\begin{figure}\small
\begin{verbatim}
{ world: { // id gives canonical URL for this module
    id: "/nell.laptop.org/chapters/alphabet-book",
    scene: "opening", // initial scene
    inherit_from: "/nell.laptop.org/core/drawing_activity",
  }, scenes: [
    { id: "opening", subject: "actors/princess",
      verb: "speak", object: "actors/user",
      text: "Hello! Would (o/2) like to hear a story?",
      group: "/nell.laptop.org/core/story-intro-group",
    },
    { id: "story-1",
      cond: rule`At(opening) && Event(Chat, Yes)`,
      subject: "actors/princess",
      verb: "speak", object: "actors/user",
      text: "Once upon a time..." },
    ...
    { id: "r-drawn",
      cond: rule`Event(Activity, DrawR)`,
      subject: "actors/princess",
      verb: "speak", object: "actors/user",
      text: "Very good, (o/3)!",
      action: js`startAnimation("r-runs")` }, ...
  ], ... }
\end{verbatim}
\caption{Excerpt of story module.  JSON syntax has been extended with
quasiquote~\cite{quasiquote}.}\label{fig:rules}
\end{figure}
To support Nell's \textit{modular narrative interface}, we use a rule-based
story model with programmable conflict resolution. (Figure~\ref{fig:rules} shows
an excerpt of the story module underlying the interaction described in
Section~\ref{sec:lesson}.)

Stories consist of a sequence of \textit{scenes} plus information about
characters, locations, and activities.  Scenes are ``goals''
in the STRIPS formalism~\cite{strips} and consist of
action descriptions, preconditions, and effects.
In principle, all preconditions for all scenes of all story modules
are evaluated continuously to determine the actions of the narrative
agents.  This production rules system
is made computationally efficient using the Rete
algorithm.  Dialog is expressed using Curveship-style
templates~\cite{montfort:curveship}.

@@

New story modules can be incorporated by visiting an internet page in
connected deployments.  In disconnected deployments, new stories
might be distributed once a year on USB sticks, or shared among
friends.  In a classroom environment, a teacher can share stories for
the day's lesson at the start of class.
